---
title: "NLP on scraped Reddit posts"
author: "Tomer & Hadas"
date: 07-15-2023
title-block-banner: "#7952B3"
execute: 
  warning: false
  message: false
  cache: true
format:
  html:
    theme: cosmo
    backgroundcolor: "#F5F5F5"
    fontcolor: "#7952B3"
    toc: true
    toc-depth: 2
    toc-location: right
    embed-resources: true
editor: visual
---

# Setup

```{r}
#| output: false
library(tidyverse)
library(quanteda)
library(quanteda.textplots)
library(text)
library(RedditExtractoR)
```

# Reddit scraping

```{r}
reddit_posts <- RedditExtractoR::find_thread_urls(subreddit = "irishpolitics", period = "month", sort_by = "hot")
rownames(reddit_posts) <- NULL

reddit_posts2 <- filter(reddit_posts, comments > 1 & comments < 7)

reddit_comment <- RedditExtractoR::get_thread_content(reddit_posts2$url)
View(reddit_comment$comments)
```

# Classifying title's sentiment

```{r}
#| eval: false
reddit_posts2$id <- c(1:nrow(reddit_posts2))

emotions <- data.frame("url" = reddit_posts2$url,
                       "Negative" = rep(NA, nrow(reddit_posts2)),
                       "Positive" = rep(NA, nrow(reddit_posts2)))

for (i in 1:nrow(reddit_posts2)) {
  tmp <- textClassify(reddit_posts2$title[i], return_all_scores = T) |>
  pivot_wider(names_from = label_x, values_from = score_x)
  
  emotions$Negative[i] <- tmp$NEGATIVE
  emotions$Positive[i] <- tmp$POSITIVE
}
```

```{r}
#| echo: false
emotions <- read_rds("../emotions.rds")
```

# Classifying comment's sentiment
## Classifying
```{r}
#| eval: false
comments <- reddit_comment$comments |>
  select(url, comment) |>
  filter(str_length(comment) <= 512)

emotions_comments <- data.frame("url" = unique(comments$url),
                                "Negative" = rep(NA, length(unique(comments$url))),
                                "Positive" = rep(NA, length(unique(comments$url))))

for (u in unique(comments$url)) {
  tmp <- textClassify(comments$comment[comments$url == u], return_all_scores = T) |>
  pivot_wider(names_from = label_x, values_from = score_x)
  
  emotions_comments$Negative[emotions_comments$url == u] <- tmp$NEGATIVE
  emotions_comments$Positive[emotions_comments$url == u] <- tmp$POSITIVE
}
```

## Comparing to post sentiment
### Renaming cols
```{r}
#| eval: false
emotions_comments <- emotions_comments |>
  rename(Negative_comment = Negative, Postive_comment = Positive)
```

```{r}
#| echo: false
emotions_comments <- read_rds("../emotions_comments.rds")
```


```{r}
combined <- emotions_comments |>
  left_join(emotions)

cor.test(combined$Negative, combined$Negative_comment)

cor.test(combined$Positive, combined$Postive_comment)
```

## Scatter plots
### Positive

```{r}
ggplot(combined, aes(x = Positive, y = Postive_comment)) +
  geom_point() +
  geom_smooth(method = "lm", color = "royalblue") +
  theme_classic()
```



```{r}
ggplot(combined, aes(x = Negative, y = Negative_comment)) +
  geom_point() +
  geom_smooth(method = "lm", color = "red") +
  theme_classic()
```


## Loading dictionary
```{r}
dict <- read.delim("shared.txt", header = F)

dict_vec <- paste(dict$V1, collapse = "|")
```

## Counting moral words in each post title
```{r}
reddit_posts2 <- reddit_posts2 |>
  mutate(n_moral = str_count(title, pattern = dict_vec))

ggplot(reddit_posts2, aes(x = n_moral, y = comments)) +
  geom_point() +
  geom_smooth(method = "lm")

parameters::model_parameters(lm(comments ~ n_moral, data = reddit_posts2))
```

## CCR analysis vs. Belief in Just world

Embedding the Just-World dictionary

```{r}
questionnaire_JW <- c("I feel that people get what they are entitled to have",
                   "I feel that a person’s efforts are noticed and rewarded",
                   "I feel that people earn the rewards and punishments they get",
                   "I feel that people who meet with misfortune have brought it on themselves",
                   "I basically feel that the world is a fair place")

questionnaire_vec <- text::textEmbed(questionnaire_JW, 
                                     model = 'sentence-transformers/all-MiniLM-L6-v2', 
                                     layers = 6)

questionnaire_vec_avg <- questionnaire_vec$texts$texts %>% 
  summarise(across(everything(), mean))
```

Extracting posts from r/Politics. This subreddit can be seen as an experimental control for r/irishpolitics.
```{r}
#| eval: false
reddit_posts_politics <- RedditExtractoR::find_thread_urls(subreddit = "Politics", period = "month", sort_by = "hot")

reddit_posts_politics <- reddit_posts_politics[1:200,]

rownames(reddit_posts_politics) <- NULL
```

```{r}
#| echo: false
reddit_posts_politics <- read_rds("../reddit_posts_politics.rds")
```


## Embedding!
```{r}
#| eval: false
politics_embedded <- textEmbed(reddit_posts_politics$title, model = 'sentence-transformers/all-MiniLM-L6-v2', 
                                     layers = 6)

irishpolitics_embedded <- textEmbed(reddit_posts$title[1:200], model = 'sentence-transformers/all-MiniLM-L6-v2', 
                                     layers = 6)
```

```{r}
#| echo: false
politics_embedded <- read_rds("../uspolitics_embedded.rds")
irishpolitics_embedded <- read_rds("../irishpolitics_embedded.rds")
```


## Cosine similarity
```{r}
cos_simil_irishpolitics <- irishpolitics_embedded$texts$texts |>
  rowwise() |>
  mutate(JW_simil = textSimilarity(across(everything()), questionnaire_vec_avg))

cos_simil_politics <- politics_embedded$texts$texts |>
  rowwise() |>
  mutate(JW_simil = textSimilarity(across(everything()), questionnaire_vec_avg))
```

```{r}
data.frame("r/irishpolitics" = cos_simil_irishpolitics$JW_simil,
           "r/politics" = cos_simil_politics$JW_simil) |>
  pivot_longer(cols = everything() ,values_to = "Cosine", names_to = "Subreddit") |>
  
  ggplot(aes(x = Subreddit, y = Cosine, color = Subreddit)) +
  geom_jitter(width = 0.05) +
  geom_boxplot()
```

```{r}
t.test(cos_simil_irishpolitics$JW_simil, cos_simil_politics$JW_simil)
```

No difference between the two subreddits. Meaning that posts are not different in their expressions of a Just-World beliefs. Also noticeable is the relative low mean cosine values. hinting that posts about general and Irish politics are not characterized by beliefs in a Just-World.




